---
title: "BFL on PHMRC: reproducing the AP results across three label-shift cases"
author: "Jason Teng, Yu Zhu, Zehang Richard Li"
output:
  pdf_document:
    toc: true
    toc_depth: 2
    latex_engine: xelatex
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", message = FALSE, warning = FALSE)
```

This example runs the paper's BFL variants on the real **PHMRC** verbal-autopsy
data, with **AP** (Andhra Pradesh) as the target and the other five sites as
sources, across the three within-target label-shift settings from the paper:

- **Case I --- no shift** (20% of AP labeled): baselines + **base**, **domain**, **partial**, **mix**.
- **Case II --- mild shift**: baselines + **domain**, **partial**, **mix**.
- **Case III --- severe shift**: baselines + **domain**, **partial**, **mix**.

`base` uses no target labels, so it is shift-invariant and only run in Case I.
The point is to show how each variant is a single **`run_BFL()`** call followed
by **`score_BFL()`** --- the only things that change are the arguments you pass.
Notice that the shift cases flip **`label_shift = TRUE`** and score on the
unlabeled subset via **`eval_idx`**; everything else is the same call. The LCVA
fitting and summary assembly live in `helpers.R`. Requires `LCVA` and `BFL`; the
PHMRC data is bundled, so run from this folder.

```{r libs}
library(LCVA)
library(BFL)
source("helpers.R")   # LCVA + PHMRC glue + per-variant input builders + baselines
```

The two **baselines** (`local_self`, `local_avg`) are LCVA models that do *not*
use the BFL package; they are shown for comparison, as in the paper. Each
`inputs_*(ctx)` assembles that variant's `local_summaries` (and `X_target`,
`Y_target`, `Y_add`) from the LCVA phis; the package calls are written out below.

# Case I: no within-target label shift

Load AP and the five sources, fit each source's local LCVA model, and split AP
into 20% labeled / 80% unlabeled. Notice that this step fitting multiple LCVA models is the most time-consuming part of the example. This step can take 20 minutes or more.

```{r prep1}
ctx1 <- prepare_ap(target  = "AP",
                   sources = c("Mexico", "Bohol", "Dar", "Pemba", "UP"),
                   seed = 1, shift = "no_shift")

c(N = length(ctx1$Y_new), labeled = length(ctx1$labeled),
  unlabeled = length(ctx1$missing), sources = length(ctx1$src_sites))
```

## Baselines: local-self and local-avg

Here we consider two simple baseline cases:
- *local-self*: train the model on the labeled subset of AP data only, and predict on the unlabeled AP data. 
- *local-avg*: train the model on each of the five sites separately, and predict on the unlabeled AP data. Then average the prediction performance.

```{r base1}
b1_self <- local_self(ctx1)
b1_avg  <- local_avg(ctx1)
round(rbind(local_self = b1_self, local_avg = b1_avg), 4)
```

## BFL-base: sources only, no target labels

In the *base* case, we use the source-domain fitted model as input to BFL, and ignore the labeled data in the target site AP.


```{r bfl-base}
inp <- inputs_base(ctx1)
fit_base <- run_BFL(inp$local_summaries, X_target = inp$X_target,
                    Y_target = inp$Y_target, label_shift = ctx1$label_shift,
                    sampler = "gibbs")
sc_base  <- score_BFL(fit_base, Y_eval = inp$Y_eval, eval_idx = inp$eval_idx)
c(top1 = sc_base$top1_acc, top1bal = sc_base$balanced_acc, csmf = sc_base$csmf_acc)
```

## BFL-domain: labels fed into a candidate model

In the *BFL-domain* case, we use the fitted models based on each source domain and the local labeled data in AP, as candidate models for BFL. That is, the labeled rows train a self-model (added to `local_summaries`).

```{r bfl-domain}
inp <- inputs_domain(ctx1)
fit_dom <- run_BFL(inp$local_summaries, X_target = inp$X_target,
                   Y_target = inp$Y_target, Y_add = inp$Y_add,
                   label_shift = ctx1$label_shift, sampler = "gibbs")
sc_dom  <- score_BFL(fit_dom, Y_eval = inp$Y_eval, eval_idx = inp$eval_idx)
c(top1 = sc_dom$top1_acc, top1bal = sc_dom$balanced_acc, csmf = sc_dom$csmf_acc)
```

## BFL-partial: labels fed into the BFL model

In the *BFL-domain* case, we use the fitted models based on each source domain as candidate models for BFL. In addition, we treat the observed labels as partially known labels in fitting the BFL step. They go straight into `Y_target` (`NA` on the unlabeled rows).

```{r bfl-partial}
inp <- inputs_partial(ctx1)
fit_par <- run_BFL(inp$local_summaries, X_target = inp$X_target,
                   Y_target = inp$Y_target, label_shift = ctx1$label_shift,
                   sampler = "gibbs")
sc_par  <- score_BFL(fit_par, Y_eval = inp$Y_eval, eval_idx = inp$eval_idx)
c(top1 = sc_par$top1_acc, top1bal = sc_par$balanced_acc, csmf = sc_par$csmf_acc)
```

## BFL-mix: a split of both

Half the labeled rows (per cause) train the self-model; the other half enter as
partial labels. Both `Y_target` and `Y_add` are supplied.

```{r bfl-mix}
inp <- inputs_mix(ctx1)
fit_mix <- run_BFL(inp$local_summaries, X_target = inp$X_target,
                   Y_target = inp$Y_target, Y_add = inp$Y_add,
                   label_shift = ctx1$label_shift, sampler = "gibbs")
sc_mix  <- score_BFL(fit_mix, Y_eval = inp$Y_eval, eval_idx = inp$eval_idx)
c(top1 = sc_mix$top1_acc, top1bal = sc_mix$balanced_acc, csmf = sc_mix$csmf_acc)
```

## Accuracy across methods (Case I)

```{r table1}
round(rbind(
  local_self = b1_self,
  local_avg  = b1_avg,
  base    = c(top1 = sc_base$top1_acc, top1bal = sc_base$balanced_acc, csmf = sc_base$csmf_acc),
  domain  = c(top1 = sc_dom$top1_acc,  top1bal = sc_dom$balanced_acc,  csmf = sc_dom$csmf_acc),
  partial = c(top1 = sc_par$top1_acc,  top1bal = sc_par$balanced_acc,  csmf = sc_par$csmf_acc),
  mix     = c(top1 = sc_mix$top1_acc,  top1bal = sc_mix$balanced_acc,  csmf = sc_mix$csmf_acc)
), 4)
```

## Inspecting a fited object

Every `run_BFL()` fit has class `BFL` and every `score_BFL()` result has class
`BFL_score`, each with `print` / `summary` / `plot`. Shown here for `domain`.

```{r s3-fit-print}
fit_dom
```

```{r s3-fit-summary}
summary(fit_dom)
```

```{r s3-fit-plot, fig.width = 7, fig.height = 4}
plot(fit_dom)
```

```{r s3-score}
sc_dom
summary(sc_dom)
```

```{r s3-score-plot, fig.width = 7, fig.height = 4}
plot(sc_dom)
```

```{r cleanup1, include = FALSE}
# Free Case I before Case II. Keep fit_base + the AP truth vector for the
# "Additional figures" (figure 8 & 10) section at the end (base, no shift).
ap_truth <- ctx1$Y_truth_origin
rm(ctx1, fit_dom, fit_par, fit_mix, inp); gc(FALSE)
```

# Case II: mild within-target label shift

`prepare_ap(shift = "mild")` builds a labeled/unlabeled split with a *mild*
prevalence shift between the two halves (a stacked frame drawn from independent
Dirichlet prevalences). The BFL calls are identical to Case I except
`label_shift = TRUE`, and scoring is on the unlabeled subset (`eval_idx`); `base`
is omitted (it is shift-invariant).

```{r prep2}
ctx2 <- prepare_ap(target  = "AP",
                   sources = c("Mexico", "Bohol", "Dar", "Pemba", "UP"),
                   seed = 1, shift = "mild")
```

## Baselines

```{r base2}
b2_self <- local_self(ctx2)
b2_avg  <- local_avg(ctx2)
round(rbind(local_self = b2_self, local_avg = b2_avg), 4)
```

## BFL fits (domain, partial, mix)

```{r bfl2}
inp    <- inputs_domain(ctx2)
sc_dom2 <- score_BFL(run_BFL(inp$local_summaries, X_target = inp$X_target,
                             Y_target = inp$Y_target, Y_add = inp$Y_add,
                             label_shift = ctx2$label_shift, sampler = "gibbs"),
                     Y_eval = inp$Y_eval, eval_idx = inp$eval_idx)

inp    <- inputs_partial(ctx2)
sc_par2 <- score_BFL(run_BFL(inp$local_summaries, X_target = inp$X_target,
                             Y_target = inp$Y_target,
                             label_shift = ctx2$label_shift, sampler = "gibbs"),
                     Y_eval = inp$Y_eval, eval_idx = inp$eval_idx)

inp    <- inputs_mix(ctx2)
sc_mix2 <- score_BFL(run_BFL(inp$local_summaries, X_target = inp$X_target,
                             Y_target = inp$Y_target, Y_add = inp$Y_add,
                             label_shift = ctx2$label_shift, sampler = "gibbs"),
                     Y_eval = inp$Y_eval, eval_idx = inp$eval_idx)
```

## Accuracy across methods (Case II)

```{r table2}
round(rbind(
  local_self = b2_self,
  local_avg  = b2_avg,
  domain  = c(top1 = sc_dom2$top1_acc, top1bal = sc_dom2$balanced_acc, csmf = sc_dom2$csmf_acc),
  partial = c(top1 = sc_par2$top1_acc, top1bal = sc_par2$balanced_acc, csmf = sc_par2$csmf_acc),
  mix     = c(top1 = sc_mix2$top1_acc, top1bal = sc_mix2$balanced_acc, csmf = sc_mix2$csmf_acc)
), 4)
```

```{r cleanup2, include = FALSE}
# Free Case II before Case III (same reason).
rm(ctx2); gc(FALSE)
```

# Case III: severe within-target label shift

`prepare_ap(shift = "severe")` draws a *severe* shift (per-cause Beta(0.2, 0.2)
unlabeled fractions, so some causes are almost entirely unlabeled). The calls are
again identical except the split, and `domain`/`mix` place the target self-model
by its self-generated (new-space) causes --- both handled inside the input
builders, so the `run_BFL()`/`score_BFL()` calls do not change.

```{r prep3}
ctx3 <- prepare_ap(target  = "AP",
                   sources = c("Mexico", "Bohol", "Dar", "Pemba", "UP"),
                   seed = 1, shift = "severe")
```

## Baselines

```{r base3}
b3_self <- local_self(ctx3)
b3_avg  <- local_avg(ctx3)
round(rbind(local_self = b3_self, local_avg = b3_avg), 4)
```

## BFL fits (domain, partial, mix)

```{r bfl3}
inp    <- inputs_domain(ctx3)
sc_dom3 <- score_BFL(run_BFL(inp$local_summaries, X_target = inp$X_target,
                             Y_target = inp$Y_target, Y_add = inp$Y_add,
                             label_shift = ctx3$label_shift, sampler = "gibbs"),
                     Y_eval = inp$Y_eval, eval_idx = inp$eval_idx)

inp    <- inputs_partial(ctx3)
sc_par3 <- score_BFL(run_BFL(inp$local_summaries, X_target = inp$X_target,
                             Y_target = inp$Y_target,
                             label_shift = ctx3$label_shift, sampler = "gibbs"),
                     Y_eval = inp$Y_eval, eval_idx = inp$eval_idx)

inp    <- inputs_mix(ctx3)
sc_mix3 <- score_BFL(run_BFL(inp$local_summaries, X_target = inp$X_target,
                             Y_target = inp$Y_target, Y_add = inp$Y_add,
                             label_shift = ctx3$label_shift, sampler = "gibbs"),
                     Y_eval = inp$Y_eval, eval_idx = inp$eval_idx)
```

## Accuracy across methods (Case III)

```{r table3}
round(rbind(
  local_self = b3_self,
  local_avg  = b3_avg,
  domain  = c(top1 = sc_dom3$top1_acc, top1bal = sc_dom3$balanced_acc, csmf = sc_dom3$csmf_acc),
  partial = c(top1 = sc_par3$top1_acc, top1bal = sc_par3$balanced_acc, csmf = sc_par3$csmf_acc),
  mix     = c(top1 = sc_mix3$top1_acc, top1bal = sc_mix3$balanced_acc, csmf = sc_mix3$csmf_acc)
), 4)
```

# Additional figures (BFL-base, no within-target label shift)

These mirror figures 8 and 10 from the CHAMPS supplement, shown here for the
**base** fit on AP under no shift. Both read straight off the fitted `BFL`
object: `fit_base$pi` (posterior cause prevalence) and `fit_base$lambda`
(posterior domain weights).

```{r fig-setup, include = FALSE}
suppressMessages(library(ggplot2))
# origin cause code i -> the i-th global PHMRC cause name
cause_names <- unique(read.csv("phmrc_clean.csv")$cause)
# true AP prevalence over the whole target (origin codes)
tp        <- table(factor(ap_truth, levels = sort(unique(ap_truth))))
codes     <- as.integer(names(tp)); true_prev <- as.numeric(tp) / sum(tp)
cause_ord <- cause_names[codes][order(true_prev)]      # y-axis order, rarest -> commonest
```

## Figure 8: estimated vs true prevalence

Black point + 95% interval = BFL-base posterior prevalence; orange star = the
true AP prevalence.

```{r fig8-base, fig.width = 7, fig.height = 6}
cc  <- as.character(fit_base$causes)                    # pi column order
grab <- function(v, code) { i <- match(as.character(code), cc); if (is.na(i)) 0 else v[i] }
pim  <- colMeans(fit_base$pi)
plo  <- apply(fit_base$pi, 2, quantile, 0.025)
phi  <- apply(fit_base$pi, 2, quantile, 0.975)
d8 <- data.frame(
  cause = factor(cause_names[codes], levels = cause_ord),
  mean  = vapply(codes, function(x) grab(pim, x), 0),
  lo    = vapply(codes, function(x) grab(plo, x), 0),
  hi    = vapply(codes, function(x) grab(phi, x), 0),
  true  = true_prev)
ggplot(d8, aes(mean, cause)) +
  geom_errorbarh(aes(xmin = lo, xmax = hi), height = 0.2) +
  geom_point(size = 2.4) +
  geom_point(aes(x = true), shape = 8, colour = "orange", size = 2.4, stroke = 1.1) +
  labs(x = "Prevalence", y = NULL,
       title = sprintf("BFL-base: estimated vs true prevalence (AP, no shift) --- CSMF %.3f",
                       sc_base$csmf_acc)) +
  theme_bw(base_size = 12)
```

## Figure 10: domain weights (lambda)

Posterior-mean weight BFL-base places on each source for each cause. This is just
the package generic `plot(fit, type = "lambda")` --- we relabel the numeric cause
codes to PHMRC cause names and pass the prevalence order so the rows match
Figure 8 (most prevalent on top).

```{r fig10-base, fig.width = 7, fig.height = 6}
f <- fit_base
f$causes <- cause_names[as.integer(f$causes)]          # origin code -> PHMRC cause name
plot(f, type = "lambda", causes = rev(cause_ord),      # rev(): most prevalent on top
     title = "BFL-base: posterior-mean domain weights (AP, no shift)")
```

# Session info

```{r session}
sessionInfo()
```
